### 2020-10-14 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
after translation of comments

It should also be noted that during the calculation process, it may be necessary to change the selected values of the feasible concessions $\delta_i$, $1 \leq i \leq s$, the concessions may be quite rigid or, conversely, excessively large. In general, the change in the order of efficiency criteria importance may be required. Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:9})
\begin{equation}\label{eq:10}
%\mathbb{P}_t=\{ P_1,P_2,\dots,P_t\},
\mathbb{F}=\{ F_1,F_2,\dots,F_t\},
\end{equation}
{\color{red} 
where $F_t, t>1,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
dynamically by adding new required MCOlex problems or removing unnecessary ones. 
The problems of the set $\mathbb{F}$ specify various requirements for the desired solutions and, thus, all problems of the set $\mathbb{F}$ should be solved in the course of calculations.
In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex)\footnote{From a general point of view, the MCOlex problem can be considered as a special case of the MMCOlex problem for $t=1$.}.
}

@@

To provide uniform calculations, a sequential scheme is applied to obtain values the constraints $g_i(y(x)) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$. According to this scheme, the constraint values are calculated strictly in their order, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme). The number of constraints, in which the values were calculated, will be referred below as the index $\nu=\nu(x)$, that is
\begin{equation}\label{eq:18}
\nu(x) : g_\nu(x)>0, g_j(x)\leq 0, 1 \leq j \leq \nu - 1, 1 \leq \nu = \nu(x) \leq m
\end{equation}
(when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index allows determining the integrated function for the problem (\ref{eq:16}) (see Fig. \ref{fig:1},
{\color{red}  the example taken from \cite{Book2013}})
\begin{equation}\label{eq:19}
F(x)=g_\nu(x), \nu=\nu(x),
\end{equation}
defined and computable everywhere in $[0,1]$. Its value at the point $x\in [0,1]$ is either the value of the left side of the constraint violated at this point (the case when $\nu \leq m$), or the value of the minimized function (the case when $\nu = m + 1$). The process of calculating the value of $F(x)$, $x\in [0,1]$, is completed either as a result of establishing the inequality $g_j(x)>0$, or as a result of reaching the value $\nu(x)=m+1$ (further in the paper, this procedure will be referred to as trial).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig1}
  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) {\color{red} initial problem functions} , b) the function $F(x)$ from (\ref{eq:19})}
  \label{fig:1}
\end{figure}

@@

Various modifications of this algorithm and the corresponding theory of convergence are presented in \cite{c18}.
{\color{red}
В частности, одна из модификаций связана с использованием информации о локальном поведении функций задачи \cite{Sergeyev2003}.
Условия сходимости алгоритма определяются следующей теоремой из \cite{c18}.
}

@@

As noted earlier, the MCOlex problem at each stage (\ref{eq:9}) of the $\varepsilon$-constraint method is a global optimization problem with nonlinear constraints (\ref{eq:16}). Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_j (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}. 
{\color{red} 
This possibility of reusing search information can be briefly presented as follows.
\begin{itemize}
\item 
При минимизации очередного критерия $f_j(y(x)), 1\leq j\leq s,$ получаемая поисковая информация накапливается в форме множества $A_k$ из (\ref{eq:21}) (следует отметить, что в $A_k$ наряду со значениями $z_i, 1\leq i\leq k,$ минимизируемого критерия $f_j(y(x))$ запоминаются и вычисленные значения векторного критерия $f(y(x_i)), 1\leq i\leq k$.
\item 
При переходе к минимизации следующего критерия $f_{j+1}(y(x))$ значения этого критерия $z_i, 1\leq i\leq k,$ во всех точках $x_i, 1\leq i\leq k,$ ранее выполненных итераций глобального поиска могут получены при помощи быстро выполнимых вычислений на основе запомненных значений векторного критерия $f(y(x_i)), 1\leq i\leq k,$ в поисковой информации $A_k$
\begin{equation}\label{new_equation}
z'_i=Z(f(y(x_i))), 1\leq i\leq k,
\end{equation}
где процедура вычисления значений $z'_i, 1\leq i\leq k$ определяется правилами (\ref{eq:9}), (\ref{eq:18})--(\ref{eq:20}), без повторных вычислениях значений критериев $f_j(y(x)), 1\leq l\leq s$.
\end{itemize}
}
This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.

Moreover, using the search information $A_k$ allows reducing the multistage solution of the MCOlex problem in accordance with the scheme (\ref{eq:9}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:9}), namely
\begin{equation}\label{eq:30}
P_{\delta lex} (f,\delta,D)=Arg \min_{y \in D} {f_s(y)}, f_j(y)\leq f_j^* + \delta_j, 1 \leq j < s,
\end{equation}
where for the unknown a priori values $f_j^*$, $1 \leq j < s$, the estimates $z_j^k$, ${1 \leq j < s}$, from (\ref{eq:23}) obtained from the search information $A_k$ can be used, and for the concessions $\delta_j$, $1 \leq j < s$, the normalized values can be applied, that is
\begin{equation}\label{eq:31}
	\begin{matrix}
	\delta_j = z_j^k + \theta_j (g_j^k-z_j^k ), 0 \leq \theta_j \leq 1, 1 \leq j < s,\\
	g_j^k = \max{\{ g_j (x_i ) : \nu(x_i)=j, 1 \leq i \leq k \}}.
	\end{matrix}
\end{equation}
{\color{red} 
При этом, как и ранее, при любых изменениях оценок  $f_j^*, \delta_j, g_j^k, 1\leq j < s, 1 \leq i \leq k,$ решение задачи (\ref{eq:30}) может быть продолжено с использованием всей имеющейся поисковой информации после быстро выполняемых вычислений новых значений $z'_i,1\leq i \leq k,$ в соответствии с процедурой (\ref{new_equation}).
}
In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.

@@

\begin{theorem}
Let the next minimized function $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previous minimized function $\phi(x)$ by some limited value $\delta$, i.e.
\begin{equation}\label{eq:32}
|\psi(x)-\phi(x)| \leq \delta, x \in [0,1],
\end{equation}
and let the stopping criterion of (\ref{eq:28}) for the specified accuracy of global search $\varepsilon>0$ be fulfilled when minimizing the function $\phi(x)$. Then 
\begin{equation}\label{eq:33}
|\psi(x^*)-\phi(x_k^*)| \leq \Delta = \frac{\alpha L \varepsilon}{2} + \delta,
\end{equation}
if starting from a certain iteration $k_0>0$ the condition is fulfilled for the estimation of the Lipschitz constant $m$ and some $\alpha > 1$
\begin{equation}\label{eq:34}
m \geq L (1+ \sqrt{\frac{\alpha + 4 \beta (1 + \beta)}{\alpha - 1}}),
\end{equation}
where
\begin{equation}\label{eq:35}
m=\max{r_l H_l^k}, 1 \leq l \leq M^k, k>k_0,
\end{equation}
\begin{equation}\label{eq:36}
L=\max{L_i}, 1 \leq i \leq s, \; L_i\; \text{from}\; (\ref{eq:3}),
\end{equation}
\begin{equation}\label{eq:beta}
\beta =  \frac{2r\delta} {L\varepsilon(r-1)},
\end{equation}
\begin{equation}\label{eq:37}
\psi(x^*)=\min{\psi(x)},x \in [0,1],
\end{equation}
\begin{equation}\label{eq:38}
x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k.
\end{equation}
\end{theorem}
The points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $H_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively.

>>>>

It should also be noted that during the calculation process, it may be necessary to change the selected values of the feasible concessions $\delta_i$, $1 \leq i \leq s$, the concessions may be quite rigid or, conversely, excessively large. In general, the change in the order of efficiency criteria importance may be required. Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:9})
\begin{equation}\label{eq:10}
%\mathbb{P}_t=\{ P_1,P_2,\dots,P_t\},
\mathbb{F}=\{ F_1,F_2,\dots,F_t\},
\end{equation}
{\color{red} 
where $F_t, \; t>1,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
dynamically by adding new required MCOlex problems or removing unnecessary ones. 
The problems of the set $\mathbb{F}$ specify various requirements for the desired solutions and, thus, all problems of the set $\mathbb{F}$ should be solved in the course of calculations.
In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex)\footnote{From a general point of view, the MCOlex problem can be considered as a special case of the MMCOlex problem for $t=1$.}.
}

@@

To provide uniform calculations, a sequential scheme is applied to obtain values the constraints $g_i(y(x)) \leq 0$, $1 \leq i \leq m$, and the minimized function $\varphi(x)$. According to this scheme, the constraint values are calculated strictly in their order, and the calculations immediately stop when the first violated constraint is detected (the partial computability scheme). The number of constraints, in which the values were calculated, will be referred below as the index $\nu=\nu(x)$, that is
\begin{equation}\label{eq:18}
\nu(x) : g_\nu(x)>0, g_j(x)\leq 0, 1 \leq j \leq \nu - 1, 1 \leq \nu = \nu(x) \leq m
\end{equation}
(when all the constraints are satisfied, $\nu=m+1$ is assumed). Introducing the index allows determining the integrated function for the problem (\ref{eq:16}) (see Fig. \ref{fig:1},
{\color{red}  the example was taken from \cite{Book2013}})
\begin{equation}\label{eq:19}
F(x)=g_\nu(x), \nu=\nu(x),
\end{equation}
defined and computable everywhere in $[0,1]$. Its value at the point $x\in [0,1]$ is either the value of the left side of the constraint violated at this point (the case when $\nu \leq m$), or the value of the minimized function (the case when $\nu = m + 1$). The process of calculating the value of $F(x)$, $x\in [0,1]$, is completed either as a result of establishing the inequality $g_j(x)>0$, or as a result of reaching the value $\nu(x)=m+1$ (further in the paper, this procedure will be referred to as trial).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig1}
  \caption{An example of the integrated function for the problem (\ref{eq:16}) using a sequential constraint calculation scheme: a) {\color{red} the initial problem functions}, b) the function $F(x)$ from (\ref{eq:19})}
  \label{fig:1}
\end{figure}

@@

Various modifications of this algorithm and the corresponding theory of convergence are presented in \cite{c18}.
{\color{red}
In particular, one modification of the algorithm involves using information about local behavior of the problem function  \cite{Sergeyev2003}.
The AGCS convergence conditions are defined by the following theorem from  \cite{c18}.
}

@@

As noted earlier, the MCOlex problem at each stage (\ref{eq:9}) of the $\varepsilon$-constraint method is a global optimization problem with nonlinear constraints (\ref{eq:16}). Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_j (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}. 
{\color{red} The search information can be reused by using the algorithm briefly described below.
\begin{itemize}
\item 
When the criterion $f_j(y(x)), 1\leq j\leq s,$ is minimized, the resulting search information is accumulated as a set $A_k$ from (\ref{eq:21}) (note that along with the values $z_i, 1\leq i\leq k,$ of the minimized criterion $f_j(y(x))$, $A_k$ also accumulates the calculated values of the vector criterion $f(y(x_i)), 1\leq i\leq k$).
\item 
When the minimization of the next criterion $f_{j+1}(y(x))$ begins, the values of this criterion $z'_i, 1\leq i\leq k,$ at all the points  $x_i, 1\leq i\leq k,$ of the previously completed iterations of the global search can be obtained via quick calculations based on the stored values of the vector criterion $f(y(x_i)), 1\leq i\leq k,$ in the search information $A_k$
\begin{equation}\label{new_equation}
z'_i=Z(f(y(x_i))), 1\leq i\leq k,
\end{equation}
where calculation procedure of the values $z'_i, 1\leq i\leq k$ is determined by the rules (\ref{eq:9}), (\ref{eq:18})--(\ref{eq:20}), without any need to repeat calculations of criteria values $f_j(y(x)), 1\leq l\leq s$.
\end{itemize}
}
This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.

Moreover, using the search information $A_k$ allows reducing the multistage solution of the MCOlex problem in accordance with the scheme (\ref{eq:9}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:9}), namely
\begin{equation}\label{eq:30}
P_{\delta lex} (f,\delta,D)=\arg \min_{y \in D} { \{ f_s(y), f_j(y)\leq f_j^* + \delta_j, 1 \leq j < s\} },
\end{equation}
where for the unknown a priori values $f_j^*$, $1 \leq j < s$, the estimates $z_j^k$, ${1 \leq j < s}$, from (\ref{eq:23}) obtained from the search information $A_k$ can be used, and for the concessions $\delta_j$, $1 \leq j < s$, the normalized values can be applied, that is
\begin{equation}\label{eq:31}
	\begin{matrix}
	\delta_j = z_j^k + \theta_j (g_j^k-z_j^k ), 0 \leq \theta_j \leq 1, 1 \leq j < s,\\
	g_j^k = \max{\{ g_j (x_i ) : \nu(x_i)=j, 1 \leq i \leq k \}}.
	\end{matrix}
\end{equation}
{\color{red} 
Considering this and previous examples, when any changes in estimates $f_j^*, \delta_j,$ $g_j^k, 1\leq j < s, 1 \leq i \leq k,$ occur, the solution of the problem (\ref{eq:30}) can be continued using all the available search information after the new values $z'_i,1\leq i \leq k,$ are promptly calculated following the procedure (\ref{new_equation}).
}
In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.

@@

\begin{theorem}
Let the next minimized function $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previous minimized function $\phi(x)$ by some limited value $\delta$, i.e.
\begin{equation}\label{eq:32}
|\psi(x)-\phi(x)| \leq \delta, x \in [0,1],
\end{equation}
and let the stopping criterion of (\ref{eq:28}) for the specified accuracy of global search $\varepsilon>0$ be fulfilled when minimizing the function $\phi(x)$. Then 
\begin{equation}\label{eq:33}
|\psi(x^*)-\phi(x_k^*)| \leq \Delta = \frac{\alpha L \varepsilon}{2} + \delta,
\end{equation}
if starting from a certain iteration $k_0>0$ the condition is fulfilled for the estimation of the Lipschitz constant $m$ and some $\alpha > 1$
\begin{equation}\label{eq:34}
m \geq L (1+ \sqrt{\frac{\alpha + 4 \beta (1 + \beta)}{\alpha - 1}}),
\end{equation}
where
\begin{equation}\label{eq:35}
m=\max{\{r_l H_l^k, 1 \leq l \leq M^k \} }, k>k_0,
\end{equation}
\begin{equation}\label{eq:36}
L=\max{\{L_i, 1 \leq i \leq s \}}, \; L_i\; \text{from}\; (\ref{eq:3}),
\end{equation}
\begin{equation}\label{eq:beta}
\beta =  \frac{2r\delta} {L\varepsilon(r-1)},
\end{equation}
\begin{equation}\label{eq:37}
\psi(x^*)=\min{\{\psi(x), x \in [0,1]\}},
\end{equation}
\begin{equation}\label{eq:38}
x_k^* = arg \min{\{\phi(x_k), 1 \leq i \leq k\}}.
\end{equation}
\end{theorem}
The points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $H_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively.
